---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# pipapi

<!-- badges: start -->
<!-- [![Codecov test coverage](https://codecov.io/gh/PIP-Technical-Team/pipapi/branch/master/graph/badge.svg)](https://codecov.io/gh/PIP-Technical-Team/pipapi?branch=master) -->
<!-- badges: end -->

The goal of pipapi is to provide a high level API to the computations and 
methods that power the Poverty and Inequality Platform (PIP).

World Bank staff who have read access to the PIP data can use the functions from
this package directly (without hitting the PIP API)

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("PIP-Technical-Team/pipapi")
```
## Getting started

The main function from the `pipapi` package is the `pip()` function. See `?pip`
for more information.

### Data access

In order to use the `pip()` you'll need to have access to PIP `data_folder`. 
The folder structure looks like this:
```{bash eval=FALSE, echo=TRUE}
data-folder-root/
├─ _aux/
│  ├─ pop-regions.fst
│  ├─ pop.fst
├─ estimations/
│  ├─ survey_means.fst
│  ├─ interpolated_means.fst
│  ├─ dist_stats.fst
├─ survey_data/
│  ├─ survey_1.fst
│  ├─ ...
│  ├─ survey_n.fst
```

```{r data-access, eval=FALSE, echo=TRUE}
# Create a list of look-up tables from the root data folder
lkups <- pipapi:::clean_api_data(
  data_folder_root = "path/to/data_folder")
```

```{r data-access-2, echo=FALSE, message=FALSE, warning=FALSE}
lkups <- pipapi:::clean_api_data(
  data_folder_root = "./tests/testdata/20210401/")

str(lkups)

```

### Run package functions
Pass the `lkups` list to the main `pip()` function to compute poverty and inequality 
statistics in your `R` session.
```{r pip}
library(pipapi)

pip(country = "AGO",
    year = 2000,
    povline = 1.9,
    lkup = lkups)
```

