---
title: "Getting started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(pipapi)
```

The `pipapi` powers the PIP API, which can be accessed at [insert URL here]().
However, if have direct access to the survey data and auxiliary datasets, you
can run PIP computation directly in your `R` session.

This vignette shows you how.

# Getting started
## Load auxiliary data
In order to use PIP computations directly in your `R` session, you will need 
access to the PIP data folder that stores all the necessary data supporting the 
API. The folder structure looks like this:
```{bash eval=FALSE, echo=TRUE}
data-folder-root/
├─ _aux/
│  ├─ pop-regions.fst
│  ├─ pop.fst
├─ estimations/
│  ├─ survey_means.fst
│  ├─ interpolated_means.fst
│  ├─ dist_stats.fst
├─ survey_data/
│  ├─ survey_1.fst
│  ├─ ...
│  ├─ survey_n.fst
```

```{r data-access, eval=FALSE, echo=TRUE}
# Create a list of look-up tables from the root data folder
lkups <- pipapi::create_lkups(
  data_dir = "path/to/data_folder")
```

```{r data-access-2, echo=FALSE, message=FALSE, warning=FALSE}
lkups <- pipapi::create_lkups(
  data_dir = "../tests/testdata/20210401/")

str(lkups)

```

## Run package functions
Pass the `lkups` list to the main `pip()` function to compute poverty and inequality 
statistics in your `R` session.
```{r pip}
pip(country = "AGO",
    year = 2000,
    povline = 1.9,
    lkup = lkups)
```

